% !TeX root = ../Thesis.tex

%*****************************************
\chapter{Background}\label{ch:relatedwork}
%*****************************************
%%\glsresetall % Resets all acronyms to not used

%***************************************************************************
\section{Distributed Databases}
A distributed database is a single logical database comprising of multiple interconnected physical database servers \cite{ramakrishnan2003database, rahimi2010distributed}. It is a collection of multiple, logically interrelated databases that are geographically dispersed but work together to provide an illusion of a single unified system, making the data distribution transparent to end users via a software system, called Distributed Database Management System (DDBMS) \cite{ozsu1999principles}. As the digital revolution ushers into a new era with Big Data in the cloud, the distributed databases aim to accommodate fault tolerance, high availability, reliability, and scalability, addressing the inefficiency and the limitations faced by traditional centralized databases when handling large amount of data \cite{bagui2015database}. \\

\noindent The primary concept of distributed databases is to split up the database and execute queries or transactions in parallel across multiple database servers \cite{ramakrishnan2003database, rahimi2010distributed}. Distributed databases utilize a shared-nothing architecture, in which each database has its own processing unit, local main memory, and storage space, operating independently from the others. In distributed database, data is spread across all the nodes, with each node employing its own Database Management System (DBMS) to manage a specific portion of data and communication between the nodes occurs via a network connection \cite{coronel2016database}. \\

\noindent Distributed databases can be broadly categorized into two types \cite{ozsu1999principles}, described as follows: 
\begin{itemize}
\item[-] \emph{Homogeneous Distributed Databases}: All database nodes in homogeneous distributed databases possess an identical database management system (DBMS) and adhere to a common global schema. This uniformity simplifies the management of the distributed databases as well as implementation of data distribution strategies. Furthermore, each node is aware of the other nodes’ existence, enabling direct communication and cooperation to process requests from users \cite{silberschatz2011database}.

\item[-] \emph{Heterogeneous Distributed Databases}: In a heterogeneous distributed database, each database node features distinct types of DDMS and operating systems. Heterogeneous distributed databases are more complex, as it requires data integration and management across different systems with varying data models, query languages, and storage structures.
\end{itemize}

\noindent Distributed databases can be further distinguished into two types, based on their architecture, which dictates how the nodes within the system interact and process data. Two common architectures of distributed database management systems (DDBMS) are as follows:

\begin{itemize}
\item[-] \emph{Client-Server / Centralized}: In this architecture, one dedicated coordinator or master database node within the cluster compiles and optimizes requests or queries and manages data across the cluster \cite{silberschatz2011database}. Meanwhile, all other nodes run and process queries in parallel. This approach ensures efficient request handling, although it may introduce a single point of failure if the master database node becomes unavailable.
\item[-] \emph{Peer-to-Peer}: This type of distributed database architecture is conceived as a collection of autonomous databases that interact, establish correspondences, and exchange query and update requests in a peer-to-peer style \cite{10.1145/1374780.1374781}. Data from the local databases can be shared among peers through user requests or queries. Each node in this architecture serves both server and client, resulting in a more resilient system with no single point of failure. However, maintaining data consistency and coordinating updates can be more challenging in a peer-to-peer architecture.
\end{itemize}

\noindent Both architectures have their advantages and drawbacks, and the choice between them depends on system requirements, scalability needs, and the desired level of fault tolerance.




%***************************************************************************
\section{Distributed Data Distribution}
\subsection{Data Partition and Sharding}
\subsection{Data Replication}
\subsection{Data Distribution}



%***************************************************************************
\section{Sharding in Distributed Databases}
A single machine, or database server, possesses a finite capacity for data storage and processing. When data volume grows too large and surpasses this capacity, the single database node becomes a bottleneck. Data sharding, a common scalability strategy, addresses this limitation by segmenting large dataset into smaller chunks, called shards, and storing them across multiple database nodes \cite{9411547, 9357711, 10.1145/2491245}. Each shard contains unique information, which enables parallel and independent processing across several database nodes. Although operating on distinct database nodes, the shards retain the original database's schema and design. By distributing the workload and reducing the risks associated with a single point of failure, this approach significantly enhances the system's performance, scalability, and resilience. \\

\noindent The implementation of a sharding system in distributed databases requires the initial determination of an appropriate sharding strategy. Database architects must consider the optimal number of shards and the methods for data distribution across multiple servers. Decisions regarding the prioritization of specific requests and the management of bulk data updates and retrieval are vital. A system characterized by frequent data modifications demands a distinct architectural approach for sharding, as opposed to the one that predominantly processes read requests. Additionally, factors such as replication, reliability, and maintenance strategy should be considered. Based on these, sharded databases can be classified into one of the following architectural types: \\

\begin{enumerate}

\item \textbf{Range-based Sharding / Dynamic Sharding} \\
This approach involves distributing data based on specified value ranges of shard keys, ensuring that each shard encompasses a continuous segment of the data. For instance, data for users with names starting from A to E could be stored on one database node, while data for users with names between F to J might be stored on another. Range-based sharding, also known as dynamic sharding, is beneficial when data can be readily partitioned based on a distinct key range. It yields good performance, if only a small number of shards are involved to process the request. However, if a request engages multiple shards and requires aggregating the results, the performance would degrade. Nevertheless, the range-scans of data from the databases is efficient \cite{ 7370983}. \\

\noindent This sharding may result in uneven data distribution and overload on a single database node, depending on the data values \cite{Harrison2021, 10.14778/2556549.2556572}. 
In the prior example, the range of values with usernames starting from A to E might have much larger entries than the usernames from F to J. Despite these potential drawbacks, this method is relatively straightforward to implement.

\item \textbf{Hash-based Sharding / Algorithmic Sharding} \\
In hash-based sharding, also referred to as algorithmic sharding, a hash function is employed to determine the database node for storing a specific data shard. The hash function accepts a key and generates a hash value (digest), and subsequently determines the data allocation node \cite{costa2015sharding, venkateswaran2017simplified}. \\

\noindent Hash-based sharding is favorable when the data lacks a natural key range, and even distribution across sites is crucial. Although this approach ensures uniform data distribution among associated database nodes, it does not shard the data according to any semantically meaningful attributes.

\item \textbf{Geo Sharding / Zone-based Sharding} \\
Geo-based sharding or zone-based sharding is a location-aware sharding approach, where data is partitioned based on a user-defined criteria that maps range of data to specific geographical regions and database nodes within those regions. The primary motivation for geo-based sharding is to place data closer in perimeter, both in terms of location and network proximity, to the users and the applications requiring that data \cite{10.1145/2433396.2433424}. Essentially, this method constitutes a range-based sharding, wherein the shard key contains geographic information and the shards themselves are geolocated. For instance, in a dataset feature a "location" field for each entry, performance and system latency can be enhanced by creating a shard for each country or region and storing the corresponding data on that shard \cite{9989965}. \\

\noindent Geo-based sharding proves particularly effective when data access patterns are predominantly based on geography. By partitioning data according to geographic regions or other location-related criteria, this approach offers advantages when it is crucial to store data closer to the users or applications that access it. Nevertheless, geographic-based sharding may lead to imbalanced data distribution across the system, as certain locations may have a higher concentration of users and data flow than others.

\item \textbf{Entity-/Relationship-/Directory-based Sharding} \\
This approach utilizes a directory service or lookup table to map data to their corresponding shards, offering flexibility in data distribution and the potential for re-sharding with minimal impact \cite{9357711}. A lookup table is a table that holds a static set of information about where specific data can be found and, in this context, it uses a shard key to track which shard holds which data. Upon receiving a request for specific data, the appropriate site, i.e., the entry for associated database node in the directory is retrieved based on the shard key. This approach is useful when the number of sites or data distribution is expected to change frequently and when related data must be consolidated within a single physical shard.


\item \textbf{Multi-level Sharding} \\
This method involves multiple levels of sharding based on data requirements, for instance, data may initially be sharded or partitioned by range, followed by the use of hashing on each key within each range \cite{chen2019sschain}. Multi-level sharding is suitable when scaling the system beyond the capabilities of a single sharding level is essential.%\cite{ChameleonDB}. 
However, it also introduces added operational complexity and computational overhead. As a result, careful design and planning are essential when implementing this approach.

\end{enumerate}

%***************************************************************************
\section{Related Works}
